Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline

### 2017-05-01 draft/ms_v5.tex
Location of figures
Final changes before submission

    The Subaru Strategic Program (SSP) is an ambitious multi-band survey
    using the Hyper Suprime-Cam (HSC) on the Subaru telescope. 
    The Wide layer of the SSP is both wide and deep, reaching a detection limit of 
    $i{\sim}26.0$ mag. 
    At these depths, it is challenging to achieve accurate, unbiased, and consistent
    photometry across all five bands. 
    The HSC data are reduced using a pipeline that builds on the prototype pipeline 
    for the Large Synoptic Survey Telescope. 
    We have developed a \texttt{Python}-based, flexible framework to inject synthetic
    galaxies into real HSC images called \synpipe{}. 
    Here we explain the design and implementation of \synpipe{} and generate 
    a sample of synthetic galaxies to examine the photometric performance of the HSC
    pipeline. 
    For stars, we achieve 1\% photometric precision at $i{\sim}19.0$ mag and 7\%
    precision at $i{\sim}25.0$ in the $i$-band (corresponding to statistical scatters 
    of ${\sim}0.01$ and ${\sim}0.05$ mag respectively).  
    For synthetic galaxies with single-\ser{} profiles, forced \cmodel{} photometry
    achieves 9\% photometric precision at $i{\sim}20.0$ mag and 13\% precision at 
    $i{\sim}25.0$ in the $i$-band (corresponding to statistical scatters of 
    ${\sim}0.15$ and ${\sim}0.22$ mag respectively). 
    We show that both \forced{} PSF and \cmodel{} photometry yield unbiased color 
    estimates that are robust to seeing conditions. 
    We identify several caveats that apply to the version of HSC pipeline used for 
    the first public HSC data release (DR1) that need to be taking into consideration. 
    First, the degree to which an object is blended with other objects impacts the 
    overall photometric performance.  
    This is especially true for point sources. 
    Highly blended objects tend to have larger photometric uncertainties, 
    systematically underestimated fluxes and slightly biased colors. 
    Second, $>20$\% of stars at $22.5< i < 25.0$ mag can be misclassified as 
    extended objects.
    Third, the current \cmodel{} algorithm tends to strongly underestimate the 
    half-light radius and ellipticity of galaxy with $i>21.5$ mag.

@@

    The overall performance of the \hscpipe{} \forced{} PSF magnitude is excellent. 
    The median $i$-band PSf magnitude precision for the \texttt{goodSeeing} \tract{} 
    is ${\sim}0.015$ mag (${\sim}1.5$\%) at $i_{\mathrm{Input}}{\sim}19.0$ mag. 
    At $i_{\mathrm{Input}}{\sim}24.0$ mag, the precision of the PSF magnitude is 
    at ${\sim}3$\% level (${\sim}0.036$ mag average difference).
    At $i_{\mathrm{Input}}{\sim}25.2$ mag, the precision decreases to
    ${\sim}7$\% with an average ${\sim}0.072$ mag difference.

    The precision of PSF magnitude for the \texttt{badSeeing} \tract{} shows similar 
    performance at $i_{\mathrm{Input}}<23.5$ mag. 
    At fainter magnitudes, the precision slightly degrades. 
    The \texttt{badSeeing} \tract{} has ${\sim}5$\% precision at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% precision at
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the precision of PSF magnitude via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds it at 1-2\% level,
    which is consistent with our results. 
    However, at fainter magnitudes, external comparisons become difficult due to the 
    lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide useful evaluations of the precision of 
    the HSC photometry down to our detection limit\footnote{Strictly speaking, the 
    precision reported here should be considered as upper limits because 
    \synpipe{} currently does not consider the systematic uncertainties in PSF 
    modeling. But this effect should be subdominant.}

    Fig \ref{fig:psf_mag} also shows that the precision of the PSF photometry is 
    not filter dependent. 
    The precision of \forced{} PSF magnitudes in $r$ and $z$ is similar to the results
    for $i$-band. 
    For $r$ band we find 1.5-4.0\% precision at $r_{\mathrm{Input}} < 24.0$ mag and 
    ${\sim}8$\% precision at $r_{\mathrm{Input}}{\sim}25.2$. 
    For $z$-band we find 1.0-5.0\% precision at $z_{\mathrm{Input}} < 24.0$ mag and 
    ${\sim}10$\% precision down to $z_{\mathrm{Input}}{\sim}25.2$ mag).

    The precision for $g$ and $y$ bands becomes slightly worse at the very faint end. 
    For $g$ band, the precision is 1.5-7.0\% at $g_{\mathrm{Input}} < 24.0$ mag and is
    ${\sim}13$\% down to $25.2$ mag.
    For $y$ band, we find 1.5-5.0\% precision at $y_{\mathrm{Input}} < 23.0$ mag and
    ${\sim}12$\% down to $24.0$ mag. 
    These differences are however consistent with the differences in the seeing 
    conditions between the filters

@@

    The PSF photometry for relatively isolated stars is accurate and unbiased in $i$ 
    and $z$ bands down to faint magnitudes. 
    For the $g$, $r$, and $y$ bands, we find mean $\Delta \mathrm{mag}_{\mathrm{PSF}}$
    that are close to the PSF flux errors estimated by \hscpipe{}.  
    \hscpipe{} tends to underestimate the fluxes of stars in these bands by 
    ${\sim}0.01-0.02$ mag. 
    The exact cause of this offset is unclear but the levels of bias is quite small 
    and not a major concern.

@@

	The precisions and accuracies of \forced{} PSF magnitudes shown here are 
	summarized in Table \ref{tab:psfmag}

@@

    Fig \ref{fig:cmodel_sn} shows that the \cmodel{} photometry displays a 
    well-behaved relation between input magnitude and output \s2n{}.  
    Compared to stars, the \s2n{} from \cmodel{} for synthetic galaxies shows a larger
    scatter at fixed input magnitude. 
    This is likely due to the fact that galaxies have a wider range in sizes and shapes 
    than point sources.  
    In $i$ band, a typical $25.2$ mag galaxy has \s2n{}${\sim}20$ from \cmodel{}, but 
    there are also galaxies at the same magnitude which have \s2n{} below 5. 
    The HSC Wide layer can detect galaxies as faint as $i{\sim}26.0$ mag but the 
    sample becomes incomplete at these faint magnitudes.
    This also introduces a selection bias if galaxies with certain structural 
    properties (e.g., more extended and lower surface brightness) are harder to detect
    than others.
    HSC survey data users who want to select flux-limited galaxy samples,
    or who intend to study the population of faint (e.g., high-$z$) galaxies should
    keep this in mind.  
    Finally, we also find that a worse seeing leads to a lower \s2n{} at fixed 
    magnitude (see Fig \ref{fig:psf_s2n} for $i$-band).

@@

    The typical precision of $i$-band flux using \forced{} \cmodel{} is at 
    ${\sim}8-10$\% level in the \texttt{goodSeeing} \tract{} at 
    $i_{\mathrm{Input}}<24.0$ mag. 
    It moderately degrades to ${\sim}13$\% at $25.0$ mag. 
    The performance is similar in $i$ band for the \texttt{badSeeing} \tract{}. 
    At $20.0 < i_{\mathrm{Input}} < 24.0$ mag, the accuracies of \forced{} \cmodel{}
    magnitudes change between \plus{}$0.017$ mag to \minus{}$0.023$ mag. 
    At $24.0 < i_{\mathrm{Input}} < 25.5$ mag, \mmag{} is around $\minus0.006$ mag,
    which suggest that \forced{} \cmodel{} photometry is unbiased down to the very 
    faint end.

@@

    \rowcolor[gray]{.85} 19.0 & \smag{} & 0.016 & 0.016  & 0.021 & 0.026 & 0.015 & 0.012 & 0.011 & 0.019 & 0.017 & 0.019 \\
    \rowcolor[gray]{.85}      & \mmag{} & 0.008 & 0.008  & 0.011 & 0.009 & 0.002 & 0.001 & 0.002 & 0.003 & 0.006 & 0.007 \\
    \hline

    \rowcolor[gray]{1.0} 20.0 & \smag{} & 0.017 & 0.018  & 0.023 & 0.029 & 0.015 & 0.013 & 0.013 & 0.020 & 0.017 & 0.019 \\
    \rowcolor[gray]{1.0}      & \mmag{} & 0.009 & 0.006  & 0.011 & 0.009 & 0.002 & 0.002 & 0.003 & 0.003 & 0.006 & 0.008 \\
    \hline

    \rowcolor[gray]{.85} 21.0 & \smag{} & 0.019 & 0.019  & 0.024 & 0.031 & 0.017 & 0.014 & 0.014 & 0.021 & 0.021 & 0.023 \\
    \rowcolor[gray]{.85}      & \mmag{} & 0.009 & 0.006  & 0.012 & 0.009 & 0.002 & 0.002 & 0.002 & 0.003 & 0.006 & 0.008 \\
    \hline

    \rowcolor[gray]{1.0} 22.0 & \smag{} & 0.023 & 0.021  & 0.025 & 0.034 & 0.018 & 0.016 & 0.016 & 0.024 & 0.033 & 0.037 \\
    \rowcolor[gray]{1.0}      & \mmag{} & 0.009 & 0.005  & 0.011 & 0.009 & 0.002 & 0.002 & 0.003 & 0.004 & 0.006 & 0.006 \\
    \hline

    \rowcolor[gray]{.85} 23.0 & \smag{} & 0.029 & 0.026  & 0.027 & 0.041 & 0.020 & 0.022 & 0.025 & 0.038 & 0.074 & 0.077 \\
    \rowcolor[gray]{.85}      & \mmag{} & 0.008 & 0.004  & 0.012 & 0.009 & 0.002 & 0.002 & 0.003 & 0.005 & 0.000 & 0.003 \\
    \hline

    \rowcolor[gray]{1.0} 24.0 & \smag{} & 0.043 & 0.039  & 0.035 & 0.054 & 0.030 & 0.043 & 0.053 & 0.071 &   0.138 &   0.153 \\
    \rowcolor[gray]{1.0}      & \mmag{} & 0.007 & 0.003  & 0.011 & 0.006 & 0.002 & 0.001 & 0.005 & 0.003 & \n0.001 & \n0.034 \\
    \hline

    \rowcolor[gray]{.85} 25.0 & \smag{} &   0.078 &   0.091  & 0.064 & 0.090 & 0.047 &   0.050 &   0.094 &   0.104 & -- & -- \\
    \rowcolor[gray]{.85}      & \mmag{} & \n0.003 & \n0.002  & 0.011 & 0.001 & 0.003 & \n0.012 & \n0.001 & \n0.010 & -- & -- \\
    \hline

>>>>

    The Subaru Strategic Program (SSP) is an ambitious multi-band survey
    using the Hyper Suprime-Cam (HSC) on the Subaru telescope. 
    The Wide layer of the SSP is both wide and deep, reaching a detection limit of 
    $i{\sim}26.0$ mag. 
    At these depths, it is challenging to achieve accurate, unbiased, and consistent
    photometry across all five bands. 
    The HSC data are reduced using a pipeline that builds on the prototype pipeline 
    for the Large Synoptic Survey Telescope. 
    We have developed a \texttt{Python}-based, flexible framework to inject synthetic
    galaxies into real HSC images called \synpipe{}. 
    Here we explain the design and implementation of \synpipe{} and generate 
    a sample of synthetic galaxies to examine the photometric performance of the HSC
    pipeline. 
    For stars, we achieve 1\% photometric precision at $i{\sim}19.0$ mag and 6\%
    precision at $i{\sim}25.0$ in the $i$-band (corresponding to statistical scatters 
    of ${\sim}0.01$ and ${\sim}0.06$ mag respectively).  
    For synthetic galaxies with single-\ser{} profiles, forced \cmodel{} photometry
    achieves 13\% photometric precision at $i{\sim}20.0$ mag and 18\% precision at 
    $i{\sim}25.0$ in the $i$-band (corresponding to statistical scatters of 
    ${\sim}0.15$ and ${\sim}0.22$ mag respectively). 
    We show that both \forced{} PSF and \cmodel{} photometry yield unbiased color 
    estimates that are robust to seeing conditions. 
    We identify several caveats that apply to the version of HSC pipeline used for 
    the first public HSC data release (DR1) that need to be taking into consideration. 
    First, the degree to which an object is blended with other objects impacts the 
    overall photometric performance.  
    This is especially true for point sources. 
    Highly blended objects tend to have larger photometric uncertainties, 
    systematically underestimated fluxes and slightly biased colors. 
    Second, $>20$\% of stars at $22.5< i < 25.0$ mag can be misclassified as 
    extended objects.
    Third, the current \cmodel{} algorithm tends to strongly underestimate the 
    half-light radius and ellipticity of galaxy with $i>21.5$ mag.

@@

    The overall performance of the \hscpipe{} \forced{} PSF magnitude is excellent. 
    The median $i$-band PSf magnitude precision for the \texttt{goodSeeing} \tract{} 
    is ${\sim}0.014$ mag (${\sim}1.3$\%) at $i_{\mathrm{Input}}{\sim}19.0$ mag. 
    At $i_{\mathrm{Input}}{\sim}24.0$ mag, the precision of the PSF magnitude is 
    at ${\sim}3$\% level (${\sim}0.030$ mag statistical scatter).
    At $i_{\mathrm{Input}}{\sim}25.0$ mag, the precision decreases to
    ${\sim}6$\% with an average ${\sim}0.062$ mag difference.

    The precision of PSF magnitude for the \texttt{badSeeing} \tract{} shows similar 
    performance at $i_{\mathrm{Input}}<23.5$ mag. 
    At fainter magnitudes, the precision slightly degrades. 
    The \texttt{badSeeing} \tract{} has ${\sim}4$\% precision at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% precision at
    $i_{\mathrm{Input}}{\sim}25.0$ mag.
    \citet{HSCDR1} evaluates the precision of PSF magnitude via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds it at 1-2\% level,
    which is consistent with our results. 
    However, at fainter magnitudes, external comparisons become difficult due to the 
    lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide useful evaluations of the precision of 
    the HSC photometry down to our detection limit\footnote{Strictly speaking, the 
    precision reported here should be considered as upper limits because 
    \synpipe{} currently does not consider the systematic uncertainties in PSF 
    modeling. But this effect should be subdominant.}

    Fig \ref{fig:psf_mag} also shows that the precision of the PSF photometry is 
    not filter dependent. 
    The precision of \forced{} PSF magnitudes in $r$ and $z$ is similar to the results
    for $i$-band. 
    For $r$ band we find 1.5-4.0\% precision at $r_{\mathrm{Input}} < 24.0$ mag and 
    ${\sim}8$\% precision at $r_{\mathrm{Input}}{\sim}25.0$. 
    For $z$-band we find 1.0-5.0\% precision at $z_{\mathrm{Input}} < 24.0$ mag and 
    ${\sim}10$\% precision down to $z_{\mathrm{Input}}{\sim}25.0$ mag).

    The precision for $g$ and $y$ bands becomes slightly worse at the very faint end. 
    For $g$ band, the precision is 1.5-7.0\% at $g_{\mathrm{Input}} < 24.0$ mag and is
    ${\sim}13$\% down to $25.0$ mag.
    For $y$ band, we find 1.5-5.0\% precision at $y_{\mathrm{Input}} < 23.0$ mag and
    ${\sim}12$\% down to $24.0$ mag. 
    These differences are however consistent with the differences in the seeing 
    conditions between the filters

@@

    The PSF photometry for relatively isolated stars is accurate and unbiased in $i$ 
    and $z$ bands down to faint magnitudes. 
    For the $g$, $r$, and $y$ bands, we find mean $\Delta \mathrm{mag}_{\mathrm{PSF}}$
    that are close to the PSF flux errors estimated by \hscpipe{}.  
    \hscpipe{} tends to underestimate the fluxes of stars in these bands by 
    ${\sim}0.01-0.02$ mag at $< 24.0$ mag. 
    The exact cause of this offset is unclear but the levels of bias is quite small 
    and not a major concern.

@@

	The precisions and accuracies of \forced{} PSF magnitudes shown here are 
	summarized in Table \ref{tab:psfmag}
%% ------------------------------------------------------------------------------------ %%

@@

    Fig \ref{fig:cmodel_sn} shows that the \cmodel{} photometry displays a 
    well-behaved relation between input magnitude and output \s2n{}.  
    Compared to stars, the \s2n{} from \cmodel{} for synthetic galaxies shows a larger
    scatter at fixed input magnitude. 
    This is likely due to the fact that galaxies have a wider range in sizes and shapes 
    than point sources.  
    In $i$ band, a typical $25.2$ mag galaxy has \s2n{}${\sim}20$ from \cmodel{}, but 
    there are also galaxies at the same magnitude which have \s2n{} below 5. 
    The HSC Wide layer can detect galaxies as faint as $i{\sim}26.0$ mag but the 
    sample becomes incomplete at these faint magnitudes.
    This also introduces a selection bias if galaxies with certain structural 
    properties (e.g., more extended and lower surface brightness) are harder to detect
    than others.
    HSC survey data users who want to select flux-limited galaxy samples,
    or who intend to study the population of faint (e.g., high-$z$) galaxies should
    keep this in mind.  
    Finally, we also find that a worse seeing leads to a lower \s2n{} at fixed 
    magnitude (see Fig \ref{fig:star_sn} for $i$-band).

@@

    The typical precision of $i$-band flux using \forced{} \cmodel{} is at 
    ${\sim}10-14$\% level in the \texttt{goodSeeing} \tract{} at 
    $i_{\mathrm{Input}}<24.0$ mag. 
    It moderately degrades to ${\sim}18$\% at $25.0$ mag. 
    The performance is similar in $i$ band for the \texttt{badSeeing} \tract{}. 
    At $20.0 < i_{\mathrm{Input}} < 24.0$ mag, the accuracies of \forced{} \cmodel{}
    magnitudes change between \plus{}$0.017$ mag to \minus{}$0.023$ mag. 
    At $24.0 < i_{\mathrm{Input}} < 25.5$ mag, \mmag{} is around $\minus0.006$ mag,
    which suggest that \forced{} \cmodel{} photometry is unbiased down to the very 
    faint end.

@@

    \rowcolor[gray]{.85} 19.0 & \smag{} & 0.017 & 0.017  & 0.021 & 0.022 & 0.014 & 0.011 & 0.011 & 0.019 & 0.016 & 0.019 \\
    \rowcolor[gray]{.85}      & \mmag{} & 0.008 & 0.006  & 0.011 & 0.009 & 0.002 & 0.001 & 0.002 & 0.003 & 0.006 & 0.007 \\
    \hline

    \rowcolor[gray]{1.0} 20.0 & \smag{} & 0.018 & 0.018  & 0.023 & 0.025 & 0.015 & 0.012 & 0.012 & 0.020 & 0.017 & 0.019 \\
    \rowcolor[gray]{1.0}      & \mmag{} & 0.009 & 0.005  & 0.011 & 0.009 & 0.002 & 0.002 & 0.003 & 0.003 & 0.006 & 0.008 \\
    \hline

    \rowcolor[gray]{.85} 21.0 & \smag{} & 0.019 & 0.018  & 0.024 & 0.026 & 0.017 & 0.014 & 0.014 & 0.021 & 0.021 & 0.022 \\
    \rowcolor[gray]{.85}      & \mmag{} & 0.009 & 0.005  & 0.012 & 0.009 & 0.002 & 0.002 & 0.002 & 0.003 & 0.004 & 0.008 \\
    \hline

    \rowcolor[gray]{1.0} 22.0 & \smag{} & 0.023 & 0.020  & 0.025 & 0.029 & 0.018 & 0.015 & 0.017 & 0.024 & 0.035 & 0.037 \\
    \rowcolor[gray]{1.0}      & \mmag{} & 0.009 & 0.005  & 0.011 & 0.009 & 0.002 & 0.002 & 0.003 & 0.004 & 0.006 & 0.006 \\
    \hline

    \rowcolor[gray]{.85} 23.0 & \smag{} & 0.031 & 0.026  & 0.029 & 0.031 & 0.021 & 0.022 & 0.025 & 0.039 &   0.076 & 0.079 \\
    \rowcolor[gray]{.85}      & \mmag{} & 0.008 & 0.004  & 0.011 & 0.009 & 0.002 & 0.002 & 0.003 & 0.005 & \n0.001 & 0.003 \\
    \hline

    \rowcolor[gray]{1.0} 24.0 & \smag{} & 0.049 & 0.038  & 0.036 & 0.046 & 0.030 & 0.046 & 0.056 & 0.082 &   0.147 &   0.146 \\
    \rowcolor[gray]{1.0}      & \mmag{} & 0.007 & 0.003  & 0.010 & 0.007 & 0.002 & 0.001 & 0.005 & 0.003 & \n0.003 & \n0.033 \\
    \hline

    \rowcolor[gray]{.85} 25.0 & \smag{} &   0.119 &   0.108  & 0.087 & 0.107 & 0.062 &   0.103 &   0.124 &   0.152 & -- & -- \\
    \rowcolor[gray]{.85}      & \mmag{} & \n0.010 & \n0.008  & 0.004 & 0.006 & 0.003 & \n0.009 & \n0.001 & \n0.010 & -- & -- \\
    \hline
